---
title: "R Notebook"
output: html_notebook
---

import libraries
```{r}
library(dplyr)
library(data.table)
library(readxl)
```


```{r}


cdf <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalie cdf.csv",
                   sep = ',',
                   header = TRUE)


sim <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf.csv",
                   sep = ',',
                   header = TRUE)

gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalie grade by game.csv",
                   sep = ',',
                   header = TRUE)



```


max games each goalie has played
```{r}
maxGames <- gradeByGame %>%
  group_by(goalie_id) %>%
  summarize("max_games" = max(game_in_career)) %>%
  merge(gNames,
        by.x = "goalie_id",
        by.y = "goalieIdForShot")

```

create table of goalie names and goalie id
```{r}
goalieName <- shotsFull %>%
  select(goalieNameForShot,goalieIdForShot) %>%
  unique


write.csv(goalieName,
          "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalieIdNames.csv")

```


goalies grade by game
```{r}

goalieGradeByGameFinal <- sim %>%
    group_by(goalie_id) %>%
    arrange(game_in_career) %>%
    mutate(diff = actual_goals - lag(actual_goals, default = first(actual_goals)),
           diff = ifelse(game_in_career == 1, actual_goals, diff)) %>%
  rename("Cumulative_Goals" = actual_goals,
         "Actual_Goals" = diff) %>% 
  arrange(goalie_id,game_in_career) %>%
  data.table()

write.csv(cdfFinal,
          "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalieGradeByGameFinal.csv")
```


createeach goalies CDF

```{r}
cdfFinal <- cdf %>%
  merge(goalieGradeByGameFinal[,c("goalie_id","game_in_career","Actual_Goals")],
        by = c("goalie_id","game_in_career")
        ) %>%
  select(-X)

write.csv(cdfFinal,
          "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/cdfFinal.csv")
```



create shots file to input into tableau
```{r}
shotsFull <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/shots_2007-2018_stats_processed.csv",
                   sep = ',',
                   header = TRUE,
                   nrows = 100)


shotsPart <-
  shotsFull %>%
  select(goalieIdForShot,
         goalieNameForShot,
         shotID,
         event,
         shotType,
         season,
         isPlayoffGame,
         goal,
         yCordAdjusted,
         xCordAdjusted,
         shotType,
         shotRebound
         ) %>%
  mutate("goal_expectation" = rnorm(nrow(shotsFull),.6,.1),
         "yCordAdjusted" = yCordAdjusted +43,
         "xCordAdjusted" = 99 - xCordAdjusted, 
         "region" = mapply(hockeyZone,xCordAdjusted,yCordAdjusted))

regionSavePercentage <- 
shotsPart %>%
  group_by(region) %>%
  summarize("goals" = sum(goal),
            "shots" = n(),
            "regionSavePercentage" = (shots-goals)/shots)

shotsUse <- shotsPart %>%
  merge(regionSavePercentage,
        by = "region")


write.csv(shotsUse,
          "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/shotsUse.csv",
          row.names = FALSE)

```




results of simulation for input into Tableau
perform file mergers in here. tableau wont create an extract
USED THIS FOR FINAL
```{r}
goalieCDF  <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_cdf_dbf(2).csv",
                   sep = ',',
                   header = TRUE) #used to be goalie cdf.csv



ggbg <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                   sep = ',',
                   header = TRUE) %>% #used to be goalie grade by game
  data.table()
gNames <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalieIdNames.csv",
                   sep = ',',
                   header = TRUE)


goalieCDFUse <- goalieCDF %>%
  merge(ggbg[,c("goalie_id","game_in_career","actual_goals")],
        by = c("goalie_id","game_in_career")) %>%
  merge(gNames,
        by.x = "goalie_id",
        by.y = "goalieIdForShot") %>%
  filter(goalieNameForShot != "")


write.csv(goalieCDFUse,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalieCDFPreJoinedUseFinal.csv",
          row.names = FALSE)

```




create file that compares players who were rookies at the start
to other players who were rookies at the start
```{r}



start <- Sys.time()

rookieFile <- read_excel("C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_analysis.xlsx",
                         sheet = "Rookies",
                         col_names = c("goalie_id","goalieName","startYr"))


gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                   sep = ',',
                   header = TRUE) %>%
  filter(goalie_id %in% rookieFile$goalie_id)



gradeShort <- gradeByGame %>% 
  filter(goalie_id != 0)

goalies <- unique(gradeShort$goalie_id)

proximity <-temp %>% 
  mutate("closeRank" =  row_number()) %>% 
  slice (0:0) 

#seq(0,nrow(gradeShort)
for (row in seq(1,nrow(gradeShort))) {
  
  goalie <- gradeShort[row,]$goalie_id
  
  numGame <- gradeShort[row,]$game_in_career

  score <- gradeShort[row,]$cdf_score
  
  temp <- gradeShort %>%
    filter(goalie_id != goalie,
           game_in_career == numGame) %>%
    mutate("score_difference" = abs(cdf_score -score ),
           "originalGoalie" = goalie,
           "isOriginal" = FALSE) %>%
    arrange((score_difference)) %>%
    slice(1:10) %>%
    mutate("closeRank" =  row_number()) %>%
    rename("similarGoalie" = goalie_id) %>%
    select(originalGoalie,similarGoalie,game_in_career,isOriginal,closeRank)

  x <- data.frame(originalGoalie = goalie,similarGoalie =goalie,game_in_career =numGame,isOriginal=TRUE,closeRank = 0)  
  
  temp <- temp %>%
    rbind(x)
  
    proximity <- proximity %>%
      rbind(temp)

  
}


end <- Sys.time()

similarAtRookie <- proximity
print(end-start)
```


```{r}



write.csv(similarAtRookie,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/proximityRookie.csv",
          row.names = FALSE)

```


table of goalies who's first game is captured in our dataset
```{r}
rookieFile <- read_excel("C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_analysis.xlsx",
                         sheet = "Rookies",
                         col_names = c("goalie_id","goalieName","startYr"))
```


create file that compares players who were not rookies at the start
to every oneother players who were rookies at the start

```{r}

start <- Sys.time()




gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                   sep = ',',
                   header = TRUE) 

nonRookie <- gradeByGame %>%
  filter(!(goalie_id  %in% rookieFile$goalie_id))  

gradeShort <- gradeByGame %>% 
  filter(goalie_id != 0)

goalies <- unique(gradeShort$goalie_id)

proximity <-temp %>% 
  mutate("closeRank" =  row_number()) %>% 
  slice (0:0) 

#seq(0,nrow(gradeShort)
for (row in seq(1,nrow(nonRookie))) {
  
  goalie <- nonRookie[row,]$goalie_id
  
  numGame <- nonRookie[row,]$game_in_career

  score <- nonRookie[row,]$cdf_score
  
  temp <- gradeShort %>%
    filter(goalie_id != goalie,
           game_in_career == numGame) %>%
    mutate("score_difference" = abs(cdf_score -score ),
           "originalGoalie" = goalie,
           "isOriginal" = FALSE) %>%
    arrange((score_difference)) %>%
    slice(1:10) %>%
    mutate("closeRank" =  row_number()) %>%
    rename("similarGoalie" = goalie_id) %>%
    select(originalGoalie,similarGoalie,game_in_career,isOriginal,closeRank)

  x <- data.frame(originalGoalie = goalie,similarGoalie =goalie,game_in_career =numGame,isOriginal=TRUE,closeRank = 0)  
  
  temp <- temp %>%
    rbind(x)
  
    proximity <- proximity %>%
      rbind(temp)

  
}


end <- Sys.time()

print(end-start)
```



```{r}
write.csv(proximity,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/proximityNonRookie.csv",
          row.names = FALSE)

```




create file that compares players who were rookies at the start
up to comparison
to other players who were rookies at the start

```{r}
gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                 sep = ',',
                 header = TRUE)

start = Sys.time()

maxGames <- gradeByGame %>%
  group_by(goalie_id) %>%
  summarize("max_games" = max(game_in_career)) 

gradeShort <- gradeByGame %>% 
  filter(goalie_id != 0,
         goalie_id %in% rookieFile$goalie_id) %>%
  merge(maxGames,
        by = "goalie_id") #%>%
  #filter(goalie_id %in% c(8476899,8470860))


proximity2 <-temp %>% 
  mutate("closeRank" =  row_number()) %>% 
  slice (0:0) 

#seq(0,nrow(gradeShort)
for (row in seq(1,nrow(gradeShort))) {
  
  #row = 4 

  goalie <- gradeShort[row,]$goalie_id
  
  numGame <- gradeShort[row,]$game_in_career

  score <- gradeShort[row,]$cdf_score
  
  origGoalie <- gradeShort %>%
    filter(goalie_id == goalie,
           game_in_career <= numGame) %>%
    select(goalie_id,game_in_career,cdf_score) %>%
    rename("orig_score" = cdf_score,
           "orig_goalie" = goalie_id)
  
  temp <- gradeShort %>%
    filter(goalie_id != goalie,
           game_in_career <= numGame,
           max_games >= numGame) %>%
    merge(origGoalie,
          by = "game_in_career") %>%
    mutate("diff" = (orig_score-cdf_score)^2) %>%
    group_by(goalie_id) %>%
    summarize("diff" = sum(diff)) %>%
    arrange(diff) %>%
    slice(1:10) %>%
    mutate("closeRank" =  row_number())  %>%
    rbind(data.frame("goalie_id" = goalie,"diff" = 0, "closeRank" = 0)) %>%
    cbind(data.frame("originalGoalie" = goalie)) %>%
    cbind(data.frame("game" = numGame)) %>%
    rename("similarGolaie" = "goalie_id",
           "difference" = "diff")
    
   
  proximity2 <- proximity2 %>%
      rbind(temp)

  
}

end = Sys.time()

print(end-start)


```

```{r}
sumSquareRookie <- proximity2

write.csv(sumSquareRookie,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sumSquareRookie.csv",
          row.names = FALSE)

```




create file that compares players who were not rookies at the start
to everyone
up to comparison

```{r}
gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                 sep = ',',
                 header = TRUE)

start = Sys.time()

maxGames <- gradeByGame %>%
  group_by(goalie_id) %>%
  summarize("max_games" = max(game_in_career)) 

gradeShort <- gradeByGame %>% 
  filter(goalie_id != 0) %>%
  merge(maxGames,
        by = "goalie_id") #%>%
  #filter(goalie_id %in% c(8476899,8470860))

nonRookie <- gradeByGame %>%
  filter(goalie_id != 0) %>%
  filter(!(goalie_id  %in% rookieFile$goalie_id))  

proximity2 <-temp %>% 
  mutate("closeRank" =  row_number()) %>% 
  slice (0:0) 

#seq(0,nrow(gradeShort)
for (row in seq(1,nrow(nonRookie))) {
  
  #row = 4 

  goalie <- nonRookie[row,]$goalie_id
  
  numGame <- nonRookie[row,]$game_in_career

  score <- nonRookie[row,]$cdf_score
  
  origGoalie <- gradeShort %>%
    filter(goalie_id == goalie,
           game_in_career <= numGame) %>%
    select(goalie_id,game_in_career,cdf_score) %>%
    rename("orig_score" = cdf_score,
           "orig_goalie" = goalie_id)
  
  temp <- gradeShort %>%
    filter(goalie_id != goalie,
           game_in_career <= numGame,
           max_games >= numGame) %>%
    merge(origGoalie,
          by = "game_in_career") %>%
    mutate("diff" = (orig_score-cdf_score)^2) %>%
    group_by(goalie_id) %>%
    summarize("diff" = sum(diff)) %>%
    arrange(diff) %>%
    slice(1:10) %>%
    mutate("closeRank" =  row_number())  %>%
    rbind(data.frame("goalie_id" = goalie,"diff" = 0, "closeRank" = 0)) %>%
    cbind(data.frame("originalGoalie" = goalie)) %>%
    cbind(data.frame("game" = numGame)) %>%
    rename("similarGolaie" = "goalie_id",
           "difference" = "diff")
    
   
  proximity2 <- proximity2 %>%
      rbind(temp)

  
}

end = Sys.time()

print(end-start)


```

```{r}
sumSquareVet <- proximity2

write.csv(sumSquareVet,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sumSquareVet.csv",
          row.names = FALSE)

```




made a pdf of simulation results from our cdf

```{r}
#gradeByGame <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
#                   sep = ',',
#                   header = TRUE)

goalieCDF  <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_cdf_dbf(2).csv",
                   sep = ',',
                   header = TRUE) #used to be goalie cdf.csv

minGame <- goalieCDF %>%
  group_by(goalie_id,game_in_career) %>%
  summarize("min_goal" = min(sim_goal_total))

pdf <- goalieCDF %>% 
  merge(minGame,
        by = c("goalie_id","game_in_career")) %>%
    group_by(goalie_id,game_in_career) %>%
    arrange(sim_goal_total) %>%
    mutate(pmf = cdf_score - lag(cdf_score, default = first(cdf_score)),
           pmf = ifelse(sim_goal_total == min_goal, cdf_score, pmf)) %>%
  arrange(goalie_id,game_in_career) %>%
  data.table()


meanGoals <- pdf %>% 
  mutate("exp" = pmf*sim_goal_total) %>%
  group_by(game_in_career,goalie_id) %>%
  summarize("mean" = sum(exp))
  

ggbg <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/sim_dbf  final.csv",
                   sep = ',',
                   header = TRUE) %>% #used to be goalie grade by game
  data.table()

gNames <- read.table(file = "C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goalieIdNames.csv",
                   sep = ',',
                   header = TRUE)


goaliePDFUse <- pdf %>%
  merge(meanGoals,
        by = c("goalie_id","game_in_career")) %>%
  merge(ggbg[,c("goalie_id","game_in_career","actual_goals")],
        by = c("goalie_id","game_in_career")) %>%
  filter(goalie_id != 0) %>%
  merge(gNames,
        by.x = "goalie_id",
        by.y = "goalieIdForShot") %>%
  filter(goalieNameForShot != "")

write.csv(goaliePDFUse,"C:/Users/bhale/OneDrive/Documents/DVA/Group Project/goaliePDFUse.csv",
          row.names = FALSE)


```



